---
title: "Review Midterm"
author: "Hermann Junhao Fan"
format: 
  html:
    embed-resources: true
toc: true
---

The questions will be the same as the sample exam, but the dataset is eggs_price data-set in lab4's folder.

```{r, echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
library(ggplot2)
library(forecast)
library(astsa) 
library(xts)
library(tseries)
library(fpp2)
library(fma)
library(lubridate)
library(tidyverse)
library(TSstudio)
library(quantmod)
library(tidyquant)
library(plotly)
library(ggplot2)
library(dplyr)
```

```{r}
egg_data <- read.csv("eggs_price_2025.csv")
colnames(egg_data) <- c("Date", "Price")
plot_ly(egg_data, x = ~as.Date(Date)) %>%
  add_lines(y = ~Price, name = 'Egg prices (USD)', line = list(color = 'blue')) %>%
  layout(title = "Time series plot in Egg Prices",
         xaxis = list(title = "Date"),
         yaxis = list(title = "Egg Prices"))
```

```{r}
ts_egg <- ts(egg_data$Price, 
                start = decimal_date(as.Date(min(egg_data$Date))), 
                frequency = 12)  #Monthly dataa


plot(ts_egg, main = "Time Series plot")
```


## Checking for stationary

```{r}
f1 <- ggAcf(ts_egg, 50) + 
  ggtitle("ACF plot without differencing") + 
  theme_minimal()

f1
```
The ACF decays slowly to 0, which is a typical sign of non-stationary. The ADF Test below indicating the same.

```{r}
tseries::adf.test(ts_egg)
```
## Lag plot:

```{r}
gglagplot(ts_egg, do.lines = FALSE) +
  ggtitle("Lag Plot of Original Egg Prices") +
  theme_minimal()
```
From lag 1-4, we are able to see a clear autocorrelation.The lag plot does not show a consistent of autocorrelation.

```{r}
require(gridExtra)

plot1<-autoplot(ts_egg, main="without the log transformation") 
plot2<-autoplot(log(ts_egg), main="log transformed data") 

grid.arrange(plot1, plot2,nrow=2)
```
There is no such big difference, so we do not need log transformation in this case.

```{r}
tseries::adf.test(ts_egg)
```
Not stationary, so take the differencing

```{r}
fig1 <- ggAcf(ts_egg, 50) +
  ggtitle("ACF of egg's price time series") + 
  theme_minimal()

fig1
```
Non-stationary

```{r}
diff <- diff(ts_egg)
fig_diff1 <- ggAcf(diff, 50) + 
  ggtitle("ACF For differenced egg price") + 
  theme_minimal()

fig_diff1
```
```{r}
tseries::adf.test(diff)
```
Now it is stationary, second differencing is not needed.

```{r}
# now we have d = 1
# for q -> ACF, acf has 2 significant spikes, so maybe q = 0, 1, 2
fig_diff_P <- ggPacf(diff, 50) + 
  ggtitle("PACF for differenced data") + 
  theme_minimal()

fig_diff_P # in this case, p = 0, 1, 2
```
```{r}

# Load necessary libraries
library(knitr)
library(kableExtra)
library(forecast)

# Create an empty matrix to store results
results_matrix <- matrix(rep(NA, 6 * 20), nrow = 20)

# Initialize index for matrix row
i <- 1

# Loop through ARIMA model parameters: d = 1,2; p = 0:4; q = 1,2
for (d in 1:2) {
  for (p in 0:2) {
    for (q in 0:2) {
      
      # Ensure 'ts_indeed' is a univariate time series by selecting the correct column
      model <- Arima(ts_egg, order = c(p, d, q), include.drift = TRUE)
      
      # Store model parameters and AIC/BIC/AICc values in matrix
      results_matrix[i, ] <- c(p, d, q, model$aic, model$bic, model$aicc)
      
      # Increment row index
      i <- i + 1
    }
  }
}

# Convert matrix to data frame
results_df <- as.data.frame(results_matrix)
colnames(results_df) <- c("p", "d", "q", "AIC", "BIC", "AICc")

# Find the row with the minimum AIC
highlight_row <- which.min(results_df$AIC)

# Generate kable table with highlighting for the row with the minimum AIC
knitr::kable(results_df, align = 'c', caption = "Comparison of ARIMA Models") %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  row_spec(highlight_row, bold = TRUE, background = "#FFFF99")  # Highlight row in yellow


```

So far, model ARIMA(2, 1, 2) is the best.

```{r}
auto.arima(ts_egg)
```
The Auto arima saying the ARIMA(4, 1, 1) is the best.

```{r}
# Loop aagin 


# Load necessary libraries
library(knitr)
library(kableExtra)
library(forecast)

# Create an empty matrix to store results
results_matrix <- matrix(rep(NA, 6 * 20), nrow = 20)

# Initialize index for matrix row
i <- 1

# Loop through ARIMA model parameters: d = 1,2; p = 0:4; q = 1,2
for (d in 1) {
  for (p in 0:4) {
    for (q in 0:2) {
      
      # Ensure 'ts_indeed' is a univariate time series by selecting the correct column
      model <- Arima(ts_egg, order = c(p, d, q), include.drift = TRUE)
      
      # Store model parameters and AIC/BIC/AICc values in matrix
      results_matrix[i, ] <- c(p, d, q, model$aic, model$bic, model$aicc)
      
      # Increment row index
      i <- i + 1
    }
  }
}

# Convert matrix to data frame
results_df <- as.data.frame(results_matrix)
colnames(results_df) <- c("p", "d", "q", "AIC", "BIC", "AICc")

# Find the row with the minimum AIC
highlight_row <- which.min(results_df$AIC)

# Generate kable table with highlighting for the row with the minimum AIC
knitr::kable(results_df, align = 'c', caption = "Comparison of ARIMA Models") %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  row_spec(highlight_row, bold = TRUE, background = "#FFFF99")  # Highlight row in yellow

```

ARIMA(4, 1, 2) is so far the best.

## Diag:

Possible models:

ARIMA(4, 1, 1) -> Auto.arima
ARIMA(2, 1, 2) -> First choice

```{r}
model_output <- capture.output(sarima(ts_egg, 2, 1, 2))

# Find the line numbers dynamically based on a keyword
start_line <- grep("Coefficients", model_output)  # Locate where coefficient details start
end_line <- length(model_output)  # Last line of output

# Print the relevant section automatically
cat(model_output[start_line:end_line], sep = "\n")
```
```{r}
model_output <- capture.output(sarima(ts_egg, 4, 1, 1))

# Find the line numbers dynamically based on a keyword
start_line <- grep("Coefficients", model_output)  # Locate where coefficient details start
end_line <- length(model_output)  # Last line of output

# Print the relevant section automatically
cat(model_output[start_line:end_line], sep = "\n")
```

Best model should be 2, 1, 2.

```{r}
fit <- Arima(ts_egg, order = c(2, 1, 2), include.drift = FALSE)
summary(fit)
```
```{r}
# Forecast the next 365 periods
forecast_result <- forecast(fit, h = 365)

# Display forecast accuracy
accuracy(forecast_result)

# Plot the forecast
autoplot(forecast_result) +
  labs(title = "ARIMA(2,1,2) Forecast",
       x = "Time",
       y = "Predicted Values") +
  theme_minimal()
```

```{r}

# Fit models individually and check residuals
f_mean <- meanf(ts_egg, h = 200)
head(f_mean$mean) #gives the forecasted values

mean(ts_egg)

f_naive <- naive(ts_egg, h = 200)
#checkresiduals(f_naive)

f_drift <- rwf(ts_egg, drift = TRUE, h = 200)
checkresiduals(f_drift)
```

```{r}

# Generate forecasts
mean_forecast <- meanf(ts_egg, h = 365)
naive_forecast <- naive(ts_egg, h = 365)
drift_forecast <- rwf(ts_egg, drift = TRUE, h = 365)
arima_forecast <- forecast(fit, h = 365)

mean_df <- data.frame(Date = time(mean_forecast$mean), Mean = as.numeric(mean_forecast$mean))
naive_df <- data.frame(Date = time(naive_forecast$mean), Naive = as.numeric(naive_forecast$mean))
drift_df <- data.frame(Date = time(drift_forecast$mean), Drift = as.numeric(drift_forecast$mean))
arima_df <- data.frame(Date = time(arima_forecast$mean), ARIMA_Fit = as.numeric(arima_forecast$mean))

ts_egg_df <- data.frame(Date = time(ts_egg), Price = as.numeric(ts_egg))

# Create Plotly plot
plot_ly() %>%
  add_lines(data = ts_egg_df, x = ~Date, y = ~Price, name = 'Original Data', line = list(color = 'black')) %>%
  add_lines(data = mean_df, x = ~Date, y = ~Mean, name = 'Mean Forecast', line = list(color = 'blue')) %>%
  add_lines(data = naive_df, x = ~Date, y = ~Naive, name = 'Naïve Forecast', line = list(color = 'red')) %>%
  add_lines(data = drift_df, x = ~Date, y = ~Drift, name = 'Drift Forecast', line = list(color = 'green')) %>%
  add_lines(data = arima_df, x = ~Date, y = ~ARIMA_Fit, name = 'ARIMA Fit', line = list(color = 'purple')) %>%
  layout(title = 'Job Postings Forecast',
         xaxis = list(title = 'Date'),
         yaxis = list(title = 'Number of Job Postings'),
         legend = list(title = list(text = 'Forecast Methods')))

```

